Data Analytics Introduction Using Python Training

### 2024-11-09 MsaPythonDataAnalytics.qmd
'updates_day3''

- Pycharm
- Spyder
- vsCode
- Rstudio
- Positron (*New in Beta*)
- Text editors
  - notepad ++
  - vim
  - sublime
- e.t.c

@@

@@

@@

>>>>

- Pycharm
- Spyder
- *jupyter*
- vsCode
- Rstudio
- Positron (*New in Beta*)
- Text editors
  - notepad ++
  - vim
  - sublime
- e.t.c

@@

Rules for variable names

* Can't start with numbers (1,2,3,4....)
* Letter, numbers, underscores are allowed in the name but ',-, spaces are not allowed

@@

## Python | Further into Hello World

### Iterations

Iterations is repeating and the most common form of iteration is `for` 
Code highligt for `for`

```{python}
#| echo: false
#| eval: false

for value in a_list:
  #do something

@@

### Intro

Pandas relies on dataframes this is excel like data format with rows/records and columns/fields. Which mankes it easy to work with.

Each row stands for an observation and columns here are variables.

data = {
  'county' : ['Nairobi','Kiambu','Kajiado','Machakos'],
  'headquarters' : ['Nairobi','Kiambu','Kajiado','Machakos'],
  'population' : [4397073,2417735,1117840,1421932]
}

df_data = pd.DataFrame(data,columns=['county','headquarters','population'])
df_data
```

Checking and investigating the dataframe

df_data.shape()
df_data.index()
df_data.columns()
df_data.info()
df_data.count()

### Pandas Summary

df_data.sum()
df_data.cumsum()
df_data.min()
df_data.max()
df_data.idmax()
df_data.idmin()
df_data.describe()
df_data.mean()
df_data.media()
```

### Pandas Import/Export [Read/Write]

**Read CSV**

```{python}
#| echo: false
#| eval: false

#import pandas as pd
df = pd.read_csv('file.csv', header = None, nrows=5)
df.to_csv("first_dataframe.csv")

**Read excel**

```{python}
#| echo: false
#| eval: false

df = pd.read_excel('excel_file.xlsx', sheet = 'Sheet1')

df_mulitple_excel = pd.ExcelFile('excel_file.xlsx')
df = pd.read_excel(df_mulitple_excel,  'Sheet1')

df.to_excel('first_dataframe.xlsx', sheet_name = 'Sheet first')
```

|Function|	Description
|-------|--------
|read_csv |	Load delimited data from a file, URL, or file-like object; use comma as default delimiter
|read_fwf |	Read data in fixed-width column format (i.e., no delimiters)
|read_clipboard|	Variation of `read_csv` that reads data from the clipboard; useful for converting tables from web pages
|read_excel|	Read tabular data from an Excel XLS or XLSX file
|read_hdf|	Read HDF5 files written by pandas
|read_html|	Read all tables found in the given HTML document
|read_json|	Read data from a JSON (JavaScript Object Notation) string representation, file, URL, or file-like object
|read_feather|	Read the Feather binary file format
|read_orc|	Read the Apache ORC binary file format
|read_parquet|	Read the Apache Parquet binary file format
|read_pickle|	Read an object stored by pandas using the Python pickle format
|read_sas|	Read a SAS dataset stored in one of the SAS system's custom storage formats
|read_spss|	Read a data file created by SPSS
|read_sql|	Read the results of a SQL query (using SQLAlchemy)
|read_sql_table|	Read a whole SQL table (using SQLAlchemy); equivalent to using a query that selects everything in that table using `read_sql`
|read_stata|	Read a dataset from Stata file format
|read_xml|	Read a table of data from an XML file
